---
title: "Homework #4"
author: "Fan Yang, fy2232"
date: "November 24, 2017"
output:
  pdf_document: default
  html_document: default
---


```{r, echo = FALSE}
set.seed(1) # Please don't remove this code!
```


# Part 1

i. 
```{r}
ahat = 2.654
xmin = 407760
f <- function(x, a = ahat, x_min = xmin){
  return ((a-1)/x_min*(x/x_min)^(-a))
}
x <- 407760:1000000
plot(x, f(x))


```


ii. 

```{r}
upper.income <- function(u, a = ahat,x_min=xmin ){
  
}


```
The mean nearly coincide with the mode of the distribution.

3. Eyeball Fit.  Now add a Beta density function to the previous plot using `stat_function()` and the density function `dbeta()`. Let `this.mean` be a variable assigned to be the mean of the `OBP` variable.  Pick parameters $\alpha = 100*this.mean$ and $\beta = 100 - \alpha$.
```{r}
this.mean <- mean(baseball$OBP)
beta_args <- list(shape1 = 100*this.mean, shape2 = 100 - 100*this.mean)
ggplot(baseball) +
  geom_histogram(aes(x = OBP, y = ..density..)) +
  geom_density(aes(x = OBP), linetype = "dashed") +
  geom_vline(xintercept = mean(baseball$OBP)) + 
  stat_function(aes(x = OBP), fun = dbeta, args = beta_args, color = "red")
```


# Part 2

(Note that at this point in the semester, Part 1 should be easy.  There are problems if you are unable to do basic exploratory data analysis and plotting.)

4. Method of moments fit. Using the values for the mean $\mu = \frac{\alpha}{\alpha + \beta}$ and variance $\sigma^2 = \frac{\alpha \beta}{(\alpha + \beta)^2(\alpha + \beta + 1)}$, solve for $\alpha$ and $\beta$ (as functions of the mean and variance).  Use these to find the Methods of Moments estimates.  With the new estimates, create a new density histogram and add this Method of Moments fit to the plot. How does it agree with the data?
```{r}
beta.MMest <- function(data) {
  #estimates \hat{\alpha} and \hat{\beta} that best fit the data
  m <- mean(data)
  v <- var(data)
  alpha <- m/v*(m*(1-m)-v)
  beta <- (1-m)/v*(m*(1-m)-v)
  return(c(alpha, beta))
}
OBP.MM   <- beta.MMest(baseball$OBP)
beta_args <- list(shape1 = OBP.MM[1], shape2 = OBP.MM[2])
ggplot(baseball) +
  geom_histogram(aes(x = OBP, y = ..density..)) +
  geom_density(aes(x = OBP), linetype = "dashed") +
  stat_function(aes(x = OBP), fun = dbeta, args = beta_args, color = "red")
```
The new plot fits the data well.

Hint: Solving for $\alpha$ and $\beta$ you should find
\[\alpha = \frac{\mu}{\sigma^2}\left[\mu(1-\mu) - \sigma^2\right], \quad \text{ and } \quad \beta = \frac{1-\mu}{\sigma^2}\left[\mu(1-\mu) - \sigma^2\right].\]
How? Solve the mean equation for $\beta$ and then plug this value of $\beta$ into the variance equation.

5. Calibration. Find the 99 percentiles of the actual distribution of the data using the `quantile()` function using `quantile(bb$OBP, probs = seq(1, 99)/100)` and plot them against the 99 percentiles of the beta distribution you just fit using `qbeta()`. How does the fit appear to you?
```{r}
quantile(baseball$OBP, probs = seq(1, 99)/100)
a <- beta_args$shape1
b <- beta_args$shape2
qqplot(baseball$OBP, qbeta((1:99)/100, shape1 = a, shape2 = b),
       ylab = "Theoretical Quantiles")
abline(0, 1, col = "red")
```
This fits well to the beta distribution.

6. Optional if you have time -- MLE fit. Create a function that calculates the (negative) log-likelihood of the distribution.  Hint: Calculate this value with code like `-sum(dbeta(your.data.here, your.alpha, your.beta, log = TRUE))`.  The function should have argument `params = c(your.alpha, your.beta)`. Find the minimum of the negative of the log-likelihood using the optimization function `nlm()`. Take the Method of Moments fit for your starting position. How do these values compare?
```{r}
neg.beta.ll <- function(params, data) {
a <- params[1]
b <- params[2]
return(-sum(dbeta(data, shape1 = a, shape2 = b, log = TRUE)))
}
beta_args
a <- beta_args$shape1
b <- beta_args$shape2
nlm(neg.beta.ll, c(a,b), data = baseball$OBP)$minimum
nlm(neg.beta.ll, c(a,b), data = baseball$OBP)$estimate
```
These two pairs of estimators are very close.


