---
title: "Practice Final Question Solutions"
author: "Cindy Rush (cgr2130) "
date: Section #001
output:
  pdf_document: default
  html_document: default
---

# Instructions (Read this completely first)

You should complete the exam by editting this file directly.  Please knit the file often, so that if you make a mistake you catch it before the end of the exam.  You will have exactly 20 minutes from the start time to complete the exam.  **At the end you must turn in your knitted .pdf file and raw .Rmd file on Courseworks.**

**When the time is up, you must shut your computer immediately.**  We will take off points from anyone whose computer is still open after time is up.

**You may use your class notes for the exam, but not the internet.  You absolutely may not communicate with anyone else during the exam.  Doing so will result in an F in this class and likely result in termination from the MA program.**  Note that your time will be tight so you will not be able to look up every bit of code from your class notes.

\newpage

# Question 1 (30 points)

Consider the following "random walk" procedure:

1. Start with $x=5$
2. Draw a random number $r$ uniformly between $-2$ and $1$.
3. Replace $x$ with $x+r$
4. Stop if $x \leq 0$
5. Else repeat


a. Write a `while()` loop to implement this procedure. Importantly, save all the positive values of $x$ that were visited in this procedure in a vector called `x.vals`, and display its entries.

```{r}
set.seed(1)

x.vals <- 5
i      <- 1
while(x.vals[i] > 0) {
  r      <- runif(1, min = -2, max = 1)
  x.vals <- c(x.vals, x.vals[i] + r)
  i      <- i+1
}
x.vals
```

b. Produce a plot of the random walk values `x.vals` from above versus the iteration number. Make sure the plot has an appropriately labeled x-axis and and y-axis.  You should plot points for the location of the random walk at each iteration and connect the points with a line.

```{r}
plot(1:length(x.vals), x.vals, type = "o", xlab = "Iteration", ylab = "Random Walk")

library(ggplot2)

walk <- data.frame(iteration = 1:length(x.vals), value = x.vals)

ggplot(walk) +
  geom_line(mapping = aes(x = iteration, y = value)) +
  geom_point(mapping = aes(x = iteration, y = value)) +
  labs(x = "Iteration", y = "Random Walk")
```

c. (8 points) Make sure you have completed the above before you begin part c.  Below is a `random.walk()` function made to implement the random walk procedure.  Unfortunately, the function has some bugs -- find them and fix them!

\begin{verbatim}
random.walk <- function(x.start = 5, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  x.vals <- x.start
  while (TRUE) {
    r <- runif(1, -2, 1)
    if (tail(x.vals + r, 1) <= 0) {
      break
    } else {
      x.vals <- c(x.vals, x.vals + r)
      print(x.vals)
    }
  }
  return(c(x.vals = x.vals, num.steps = length(x.vals)))
}

# random.walk(x.start = 5, seed = 3)$num.steps # Should print 8
# random.walk(x.start = 10, seed = 7)$num.steps # Should print 14
\end{verbatim}

```{r}
random.walk <- function(x.start = 5, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  x.vals <- x.start
  while (TRUE) {
    r <- runif(1, -2, 1)
    if (tail(x.vals + r, 1) <= 0) {
      break
    } else {
      x.vals <- c(x.vals, x.vals[length(x.vals)] + r)
    }  
  }
  return(list(x.vals = x.vals, num.steps = length(x.vals)))
}
random.walk(x.start = 5, seed = 3)$num.steps # Should print 8
random.walk(x.start = 10, seed = 7)$num.steps # Should print 14
```

