---
title: 'Homework #2'
author: "Fan Yang (fy2232)"
date: "November 14, 2017"
output:
  pdf_document: default
  html_document: default
---

i.
```{r}
schedual <- readLines("NetsSchedule.html", warn = FALSE)
length(schedual)
sum(nchar(schedual))
max(nchar(schedual))
```
a. There are 828 lines.
b. Total number of characters is 129188.
c. The maximum number of characters is 9736.

ii.

They playing with Indiana first at 7:00 PM, Wed, Oct 18.
They playing with Boston last at 8:00 PM, Wed, Apr 11.

iii.
The $321^{st}$ line in the file holds information about the first game of the regular season.
The $402^{st}$ line in the file provides the date, time, and opponent for the final game.

iv.
```{r}
regdate <- ">[A-Z].{2},.*[0-9]<"
length(grep(regdate, schedual))
grep(regdate, schedual)
```

v.
```{r}
rdate       <- regexpr(regdate, schedual)
date <- regmatches(schedual, rdate)
date <- substring(date, 2, nchar(date) - 1)
head(date)
```

vi.

```{r}
regtime <- "<td>[0-9].*M</td>"
rtime       <- regexpr(regtime, schedual)
time <- regmatches(schedual, rtime)
time <- substring(time, 5, nchar(time) - 5)
head(time)
```

vii.

```{r}
reghome <- "<li\\sclass=\"game-status\">."
rhome      <- regexpr(reghome, schedual)
home <- regmatches(schedual, rhome)
home <- gsub("<li\\sclass=\"game-status\">@","1",home)
home <- gsub("<li\\sclass=\"game-status\">v","0",home)
home <- as.numeric(home)
sum(home)
```

viii.

```{r}
regopponent <- "class=\"team-name\">.*>[A-Z]"
ropponent      <- regexpr(regopponent, schedual)
opponent <- regmatches(schedual, ropponent)
opponent <- gsub("class.*[a-z]/","",opponent)
opponent <- gsub("\">.*","",opponent)
head(opponent,10)
```
ix.

```{r}
nets <- data.frame(date,time,opponent,home)
head(nets,10)
```
The results match the first 10 games as website.