---
title: 'Homework #1'
author: "Fan Yang (fy2232)"
date: "September 26, 2017"
output:
  pdf_document: default
  html_document: default
---

Part 1: Loading and Cleaning the Data in R
i.

```{r}
setwd("D:/Subject Materials/Columbia/STAT COMP & INTRO TO DATA SCI/HW")
housing <- read.csv('properties.csv')
```

ii.

```{r}
ncol(housing)
nrow(housing)
```

iii.
```{r}
apply(is.na(housing), 2, sum)
```

This command gives a table of the NA values in each column.

iv.
```{r}
vec <- which(housing$assessed_value!=0)
housing <- housing[vec,]
```

v.
```{r}
16319-nrow(housing)
```

vi.

```{r}
logValue <- c()
logValue <- log(housing$assessed_value)
housing <- cbind(housing, logValue)
min(logValue)
median(logValue)
mean(logValue)
max(logValue)
```

vii.

```{r}
logUnits <- log(housing$res_units)
housing <- cbind(housing, logUnits)
```

viii.

```{r}
housing <- cbind(housing, after2000 = housing$year_built>=2000)
```

Part 2: EDA

i.

```{r}
plot(housing$logUnits,housing$logValue,xlab="logUnits",ylab="logValue",cex=0.5)
```

ii.

```{r}
plot(housing$logUnits,housing$logValue,
     col = factor(housing$after2000),
     xlab="logUnits",ylab="logValue",cex=0.5)
cov(housing$logUnits,housing$logValue,use="pairwise.complete.obs")
```

The variation between the two variables is 1.504415. 
In the plot we can see that the red points are almost above the black points at the same x.
Which tells us that the value of housing built after 2000 is greater than that before 2000.
And with the units in the property increase, the value also increase.

iii.
  (i)the whole data,
```{r}
cor(housing$logUnits,housing$logValue,use="pairwise.complete.obs")
```
  (ii) just Manhattan
```{r}
cor(housing$logUnits[housing$boro_name=='Manhattan'],
    housing$logValue[housing$boro_name=='Manhattan'],
    use="pairwise.complete.obs")
```
  (iii) just Brooklyn
```{r}
cor(housing$logUnits[housing$boro_name=='Brooklyn'],
    housing$logValue[housing$boro_name=='Brooklyn'],
    use="pairwise.complete.obs")
```
  (iv) for properties built after 2000
```{r}
cor(housing$logUnits[housing$after2000],
    housing$logValue[housing$after2000],
    use="pairwise.complete.obs")
```
  (v) for properties built before 2000
```{r}
cor(housing$logUnits[housing$year_built<2000],
    housing$logValue[housing$year_built<2000],
    use="pairwise.complete.obs")
```  

iv.
```{r}
plot(range(housing$logUnits, na.rm = TRUE),
     range(housing$logValue, na.rm = TRUE),
     xlab= "logUnits", ylab = "logValue",type='n')

lines(housing$logUnits[housing$boro_name=='Manhattan'],
      housing$logValue[housing$boro_name=='Manhattan'],
      type='p',col="blue",cex=0.5)
lines(housing$logUnits[housing$boro_name=='Brooklyn'],
      housing$logValue[housing$boro_name=='Brooklyn'],
      type='p',col="red",cex=0.5)
legend("topleft",c("Manhattan","Brooklyn"),col=c("blue","red"),pch=1)
```

v.
```{r}
med.value <- median(housing$assessed_value[housing$boro_name=='Manhattan'], na.rm = TRUE)
```

vi.
```{r}
boxplot(housing$logValue ~ housing$boro_name, 
        ylab = "logValue", xlab = "boroughs")
```

vi.
```{r}
tapply(housing$assessed_value,housing$boro_name,median)
```

  