---
title: "Lab 8"
author: "Fan Yang (fy2232)"
date: "December 05, 2016"
output:
  pdf_document: default
  html_document: default
---


1. Load the dataframes `sprint.m.fastest.csv` and `sprint.w.fastest.csv`.
```{r}
sprint.m.fastest <- read.csv("sprint.m.fastest.csv", header = T, as.is = T)
sprint.w.fastest <- read.csv("sprint.w.fastest.csv", header = T, as.is = T)
```

2. First find the common track meets between the two data frames, i.e. the common entries in `CityDate`. 
```{r}
common.meets <- intersect(sprint.m.fastest$CityDate,sprint.w.fastest$CityDate)
```

3. Now compute the rows of each dataframe that correspond to these common track meets.
```{r}
ind.m <- which(is.element(sprint.m.fastest$CityDate,common.meets))
ind.w <- which(is.element(sprint.w.fastest$CityDate,common.meets))
```

4.   Now create a new dataframe that merges the columns of `sprint.m.fastest` with `sprint.f.fastest` and arrange it so that the dataframe only has three columns: `MensTime`, `WomensTime`, and `CityDate` (the common track meet).  Display the first five rows.
```{r}
sprint <- cbind(MensTime = sprint.m.fastest$Time[ind.m], 
                WomensTime = sprint.w.fastest$Time[ind.w], 
                CityDate = common.meets)
head(sprint,5)

all(sprint.m.fastest$CityDate[ind.m]==sprint.w.fastest$CityDate[ind.w])
```

5. Get the same result as 1-4 using `merge()`.
```{r}
sprint2 <- merge(sprint.m.fastest,sprint.w.fastest,by="CityDate")
sprint2 <- data.frame(MensTime = sprint2$Time.x,
                      WomensTime = sprint2$Time.y,
                      sprint2$CityDate)
all(sprint2 == sprint)
```

