---
output:
  pdf_document: default
  html_document: default
---

## import data


```{r}
library(reshape2)
library(ggplot2)
```


```{r}
crime <- read.table('APPENC02(1).txt')
colnames(crime) <- c("id","county","state","area","population",
                     "perc.young","perc.old","physicians",
                     "hospital.beds","n.crimes","perc.hs",
                     "perc.bs","perc.poor","unemployment",
                     "per.income","tot.income","region")
crime['pop.density'] = crime$population/crime$area
crime['physician.per.1000'] = crime$physicians/(crime$population/1000)
crime['beds.per.1000'] = crime$hospital.beds/(crime$population/1000)
crime['crime.rate.per.1000'] = crime$n.crimes/(crime$population/1000)
head(crime)
dim(crime)
```


```{r}
lm.crime.full <- lm(crime.rate.per.1000 ~ area + perc.young + perc.old +
                    hospital.beds + perc.bs + perc.poor + unemployment +
                    per.income + I(region) + pop.density + physician.per.1000 + 
                    beds.per.1000, data = crime)
summary(lm.crime.full)
step(lm.crime.full,direction = "backward")$anova
```



```{r}
lm.crime.fit1 <- lm(crime.rate.per.1000 ~ perc.young + hospital.beds + perc.poor + per.income +
                    I(region) + pop.density + beds.per.1000, data = crime)
summary(lm.crime.fit1)
```



```{r}
step(lm.crime.fit1,direction = "backward")$anova
```



# Poisson regression 


```{r}
df <- crime
```


```{r}
# correlation map
UNI <- 2823
set.seed(2823)
index <- sample(c(1:440))
train_df <- df[index[1:300],]
numeric_features <- c('area','population','perc.young','perc.old',
                      'physicians','hospital.beds','n.crimes','perc.hs',
                      'perc.bs','perc.poor','unemployment','per.income',
                      'tot.income','pop.density','physician.per.1000',
                      'beds.per.1000','crime.rate.per.1000')
cormat <- round(cor(train_df[numeric_features]),2)

melted_cormat <- melt(cormat)

options(repr.plot.width=12, repr.plot.height=5)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name="Pearson\nCorrelation") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1,
                                   size = 12, hjust = 1))+
  coord_fixed()
```


### train a fisrt poisson model


```{r}
#poission 

m2 <- glm( n.crimes/(population/1000)~ perc.young + perc.poor + per.income +
          factor(region) + log(pop.density) + physician.per.1000 + beds.per.1000 +
          perc.bs + unemployment, family=quasipoisson, data=df, weights=(population/1000))
summary(m2)
```




### detect outlier


```{r}
#h=0.25
outliers <- which(hatvalues(m2)>0.25)
a1<-hatvalues(m2)
a2<-residuals(m2)
a3<-data.frame(hat.values=a1,res=a2)

options(repr.plot.width=5, repr.plot.height=4)
ggplot(data=a3,aes(hat.values,res)) +
  geom_point(alpha=0.8)+
  geom_text(data=a3[outliers,],aes(hat.values-0.005,res, label=df$county[outliers]),size=3,hjust=1,alpha=0.8) +
  xlab("hatvalues") + ylab("residuals")
```




### delete outlier and retrain model


```{r}
df2 = df[-outliers,]

#poission 

m3 <- glm( n.crimes/(population/1000)~  perc.poor + per.income +
          factor(region) + log(pop.density) + physician.per.1000 + beds.per.1000 ,
          family=quasipoisson, data=df, weights=(population/1000))
summary(m3)
m3$coefficients
exp(m3$coefficients)
```


### retrain model
delete `per.income` and `physician.per.1000`


```{r}
#poission 

m4 <- glm( n.crimes/(population/1000)~ perc.poor +
          factor(region) + log(pop.density) + beds.per.1000,
          family=quasipoisson, data=df2, weights=(population/1000))
summary(m4)
```


```{r}
#h=0.25
outliers <- which(hatvalues(m4)>0.25)
a1<-hatvalues(m4)
a2<-residuals(m4)
a3<-data.frame(hat.values=a1,res=a2)

options(repr.plot.width=5, repr.plot.height=4)
ggplot(data=a3,aes(hat.values,res)) +
  geom_point(alpha=0.8)+
  geom_text(data=a3[outliers,],aes(hat.values,res, label=df$county[outliers]),size=3,hjust=1,alpha=0.8)
```


## EDA
```{r}
plot(crime$crime.rate.per.1000)

# logi.data <- crime
# logi.data$crime.rate.level <- logi.data$crime.rate.per.1000
# logi.data$crime.rate.level[which(crime$crime.rate.per.1000 <= 30)] <- 1
# logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 30 & 
#                                    crime$crime.rate.per.1000 <= 60)] <- 2
# logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 60 & 
#                                    crime$crime.rate.per.1000 <= 90)] <- 3
# logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 90 & 
#                                    crime$crime.rate.per.1000 <= 120)] <- 4
# logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 120)] <- 5

logi.data <- crime
logi.data$crime.rate.level <- logi.data$crime.rate.per.1000
logi.data$crime.rate.level[which(crime$crime.rate.per.1000 <= 60)] <- 1
logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 60 & 
                                   crime$crime.rate.per.1000 <= 120)] <- 2
logi.data$crime.rate.level[which(crime$crime.rate.per.1000 > 120)] <- 3
```

```{r}

library(MASS)
polr.fit1 <- polr(factor(crime.rate.level) ~ perc.young + hospital.beds + perc.poor + per.income +
                    I(region) + pop.density + beds.per.1000, data =logi.data)
summary(polr.fit1)
1-pchisq(deviance(polr.fit1),df.residual(polr.fit1))
polr.fit2 <- polr(factor(crime.rate.level) ~ perc.young + hospital.beds + perc.poor +
                     pop.density + beds.per.1000, data =logi.data)
summary(polr.fit2)
1-pchisq(deviance(polr.fit2),df.residual(polr.fit2))
drop1(polr.fit2,test = "Chi")

```


```{r}
library(VGAM)
logit.fit1 <- vglm(crime.rate.level ~ perc.young + hospital.beds + perc.poor + per.income +
                    I(region) + pop.density + beds.per.1000, data =logi.data,
             family=cumulative(parallel=TRUE))
summary(logit.fit1)

# logit.fit2 <- vglm(crime.rate.level ~ perc.young + hospital.beds + perc.poor + per.income +
                    # I(region) + pop.density + beds.per.1000, data =logi.data, family=cumulative)
# summary(fit2)
# pchisq(deviance(fit1)-deviance(fit), df=df.residual(fit1)-df.residual(fit),
# lower.tail=FALSE)


# step(logit.fit2,direction = "backward")$anova
```

Our main goal is to find the cause of high crime rate. Since crime rate is a quantitative variable, we simply divide this variable into 5 groups, which are `very low`(<30), `low`(30-60), `medium`(60-90), `high`(90-120) and `very high`(>120).

```{r}
table(logi.data$crime.rate.level)
# labels <- c("<30", "30-60", "60-90", "90-120",">120")
labels <- c("<60\u2030", "60\u2030-120\u2030", ">120\u2030")
x = table(logi.data$crime.rate.level)
piepercent <- round(100*x/sum(x), 1)
pie(x,labels=paste(piepercent,"%"),main = "num of crime per 1000 population",col =rainbow(length(x)))
legend("topright", labels, cex = 0.8,
   fill = rainbow(length(x)))

ind.region1 <- which(logi.data$region == 1)
ind.region2 <- which(logi.data$region == 2)
ind.region3 <- which(logi.data$region == 3)
ind.region4 <- which(logi.data$region == 4)

x = table(logi.data$crime.rate.level[ind.region1])
piepercent <- round(100*x/sum(x), 1)
pie(x,labels=paste(piepercent,"%"),main = "Crime rate in NE",col =c(7,5,2))
legend("topright", labels, cex = 0.8,
   fill = c(7,5,2))

x = table(logi.data$crime.rate.level[ind.region2])
piepercent <- round(100*x/sum(x), 1)
pie(x,labels=paste(piepercent,"%"),main = "Crime rate in NC",col =c(7,5,2))
legend("topright", labels, cex = 0.8,
   fill = c(7,5,2))

x = table(logi.data$crime.rate.level[ind.region3])
piepercent <- round(100*x/sum(x), 1)
pie(x,labels=paste(piepercent,"%"),main = "Crime rate in S",col =c(7,5,2))
legend("topright", labels, cex = 0.8,
   fill = c(7,5,2))

x = table(logi.data$crime.rate.level[ind.region4])
piepercent <- round(100*x/sum(x), 1)
pie(x,labels=paste(piepercent,"%"),main = "Crime rate in W",col = c(7,5))
legend("topright", labels[1:2], cex = 0.8,
   fill = c(7,5))


```
Then we fit a Cumulative logit model to find the factor that influence the crime rate
```{r}
names(logi.data)
library(VGAM)
logit.fit1 <- vglm(crime.rate.level ~ perc.young + perc.old + perc.poor +
                      perc.hs  + per.income + physician.per.1000 + beds.per.1000 +
                    I(region) + pop.density , data =logi.data,
             family=cumulative(parallel=TRUE))
summary(logit.fit1)
```


```{r}
names(logi.data)
library(MASS)
polr.fit1 <- polr(factor(crime.rate.level) ~ perc.young + perc.old + perc.poor +
                      perc.hs + perc.bs + per.income + physician.per.1000 + beds.per.1000 +
                    I(region) + pop.density , data =logi.data)
summary(polr.fit1)
```
AIC = 864.7338 `perc.old` `perc.hs` `perc.bs` and `physician.per.1000`

After fit the logistic model, I found the most of the predictor variables are significant. But the `perc.old` `perc.hs` `perc.bs` and `physician.per.1000` are not so important. So I decide to drop those 4 variables. In order to provide a solid evidence for dropping variables. I used AIC criterion. For the current full model, the AIC is 864.7338. Then I used backstep method to select variables.


```{r}
polr.fit2 <- polr(factor(crime.rate.level) ~ perc.young  + perc.poor +
                      per.income  + beds.per.1000 +
                    I(region) + pop.density , data =logi.data)
summary(polr.fit2)

logit.fit2 <- vglm(crime.rate.level ~ perc.young  + perc.poor +
                      per.income  + beds.per.1000 +
                    I(region) + pop.density , data =logi.data,
             family=cumulative)
summary(logit.fit2)
logit.fit2@coefficients
exp(logit.fit2@coefficients)
```

When we drop `perc.old` `perc.hs` `perc.bs` and `physician.per.1000`, the AIC now has been reduced to 859.7049, which is the least among all models. Therefore we can drop those four variables and get our final model.

Till now we have selected the most important variables `perc.young` `perc.poor` `per.income` `beds.per.1000` `pop.density` and `region`. The model now is 
$$
\log\left( \frac{P(Y\leq j)}{1-P(Y\leq j)}\right) =  \beta_j - 0.1045*perc.young \\
- 0.1909*perc.poor -1.063e-04* per.income\\
- 0.3215*beds.per - 0.9111*I(region) \\
-2.343e-04*pop.density\\
$$

```{r}
# stepAIC(logit.fit2,direction="backward",trace=FALSE)
# install.packages("usdm")
library(usdm)

vif.df = data.frame(perc.young = logi.data$perc.young, perc.old = logi.data$perc.old,
                    perc.poor = logi.data$perc.poor, perc.hs = logi.data$perc.hs,
                    perc.bs = logi.data$perc.bs, per.income = logi.data$per.income,
                    physician.per.1000 = logi.data$physician.per.1000,
                    beds.per.1000 = logi.data$beds.per.1000,
                    pop.density = logi.data$pop.density,
                    unemployment = logi.data$unemployment,
                    area = logi.data$area)

vif(vif.df)
# ggplot() + hist(vif(vif.df)$VIF)
pairs(vif.df)

vif.num <- as.vector(vif(vif.df)$VIF)
vif.labels <- c('percent young','percent old','percent poor','percent hs','percent bs','percent income',
                'physician per 1000','beds per 1000','population density', 'unemployment', 'area')
names(vif.num) <- vif.labels

ggplot() + 
  geom_point(aes(x=vif.labels,y=vif.num) ) +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = 'percent young',y=2.3, label='percent young'),size=3.5) + 
  geom_text(aes(x = 'percent old',y=2.3, label='percent old'),size=3.5) +
  geom_text(aes(x = 'percent poor',y=3.9, label='percent poor'),size=3.5) +
  geom_text(aes(x = 'percent hs',y=3.7, label='percent hs'),size=3.5) +
  geom_text(aes(x = 'percent bs',y=6.3, label='percent bs'),size=3.5,col='red') +
  geom_text(aes(x = 'percent income',y=4.3, label='percent income'),size=3.5) +
  geom_text(aes(x = 'physician per 1000',y=3.5, label='physician per 1000'),size=3.5) +
  geom_text(aes(x = 'beds per 1000',y=2.3, label='beds per 1000'),size=3.5) +
  geom_text(aes(x = 'population density',y=1.5, label='population density'),size=3.5) +
  geom_text(aes(x = 'unemployment',y=2.3, label='unemployment'),size=3.5) +
  geom_text(aes(x = 'area',y=1.5, label='area'),size=3.5) +
  geom_abline(intercept = 5, slope = 0,col='red') +
  xlab('variables') + ylab('VIF') +
  geom_point(aes(x = 'percent bs',y=max(vif.num), label='percent bs'),size=2,col='red')
  

```
```{r}
ggplot() + 
  geom_histogram( aes(crime$perc.young),
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="percent of young") +
  labs(x="percent", y="Count") +
  geom_text(aes(x = crime$perc.young[6], y=7, label='Kings'),size=5) 


ggplot() + 
  geom_histogram( aes(crime$perc.poor),
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  # labs(title="percent of poor") +
  labs(x="percent of poor", y="Count") +
  geom_text(aes(x = crime$perc.young[6], y=7, label='Kings'),size=6) 

ggplot() + 
  geom_histogram( aes(crime$beds.per.1000),
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="hospital beds per 1000 people") +
  labs(x="percent", y="Count") +
  geom_text(aes(x = crime$beds.per.1000[6], y=7, label='Kings'),size=5) 

ggplot() + 
  geom_histogram( aes(crime$pop.density),
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  # labs(title="population density") +
  labs(x="population density", y="Count") +
  geom_text(aes(x = crime$pop.density[6], y=20, label='Kings'),size=6) 
```



